---
title: "tp2"
author: "me"
date: "2023-01-31"
output: pdf_document
---

# Statistics for Biology 2: TP 2
## Exercise 2
1. Upload tauber.csv. Assign column height to variable H. How many children in the sample are taller than 110 cm? How many have height 110 or less?

```{r}
tauber <- read.table("tauber.csv", header = TRUE, sep = ";")
summary(tauber)
```
```{r}
H <- tauber$height
length(H[H>110])
```
```{r}
length(H[H<=110])
```
2. Plot cumulated frequencies for the event H>110, with ordinates in the interval (0,1): plot option ylim=c(0,1).
```{r}
cumfreq <- cumsum(H>110)/length(H)
plot(cumfreq, ylim=c(0,1), type="l")
```
3. Assign to vector rH a random permutation of H. Superpose on the same graphics cumulated frequencies for rH>110, in blue.

```{r}
set.seed(420)
rH <- sample(H)
plot(cumfreq, ylim=c(0,1), type="l", col="red", ylab = "Cumulative frequency of H > 110")
lines(cumsum(rH>110)/length(rH), col="blue")
```
4. Assign to vector iH the values of H, sorted in increasing order. Superpose on the same graphics cumulated frequencies for iH>110, in green. In which interval is the green curve constant?
```{r}
iH <- sort(H)
plot(cumfreq, ylim=c(0,1), type="l", col="red", ylab = "Cumulative frequency of H > 110")
lines(cumsum(rH>110)/length(rH), col="blue")
lines(cumsum(iH>110)/length(iH), col="green")
```
5. Assign to vector dH the values of H, sorted in decreasing order. Superpose on the same graphics cumulated frequencies for dH>110, in red. In which interval is the red curve constant ?
```{r}
dH <- sort(H, decreasing=TRUE)
plot(cumfreq, ylim=c(0,1), type="l", col=1, ylab = "Cumulative frequency of H > 110", main="Cumulative frequencies")
lines(cumsum(rH>110)/length(rH), col=2)
lines(cumsum(iH>110)/length(iH), col=3)
lines(cumsum(dH>110)/length(dH), col=4)
legend(x=0, y=0.98, legend=c("H", "rH", "iH", "dH"), col=1:4, lwd=2)
```
## Exercise 3 . Import and export of external data

Import the imcenfant.txt file using the read.table() function. Pay attention to the header, sep, dec, row.names parameters.
```{r}
D <- read.table("imcenfant.txt", header = TRUE, dec=",")
D$SEXE <- factor(D$SEXE)
D$zep <- factor(D$zep)
```

1.	Display the contents of this object, which will be named D.
```{r}
View(D)
summary(D)
```

2.	Verify the names of the columns of D. Name the rows.
```{r}
colnames(D)
```
```{r}
rownames(D) <- 1:length(D[,1])
```

3.	Add a IMC column to D. Extract from D the lines for children with IMC<15 and age<=3.5. Give the number of childrens checking the above conditions.
```{r}
D$IMC <- D$poids/(D$taille*0.01)^2
D$age <- D$an + D$mois*0.1
D[D$IMC>15 & D$age<=3.5,]
```

4.	Export the new data table to your directory as a text file.
```{r}
write.table(D, "new_imcenfant.txt")
```


## Exercise 4  Data analysis

1.	Save in your working directory the Poids_naissance.txt  file available in AmeTICE

2.	Import this  data set into R, creating an object that will be of data.frame type, named E. To do this, use the read.table() function  by correctly filling the parameters  header, sep, dec, row.names

3.	Add a PTL1 (number of preterm antecedents) variable with three modalities (where the third will be rated 2 and will correspond to 2 or more preterm antecedents).

4.	Same question for FVT (number of visits to a doctor) to add FVT1.

5.	Order the data frame according to increasing birth weights (BWT)

6.	Extract individuals with black or white mothers who smoke.



